Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-23 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

The experiment is a network of twelve 6.25ha plots that underwent a
gradient of three logging, thinning and fuelwood cutting treatments
(Table \ref{tab:Tab1}) attributed according to a randomized plot design
with three replicate blocks of four plots. The disturbance corresponds
to averages of 10 trees removed per hectare with a diameter at 1.3 m
height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
(T3). Treatments T2 and T3 besides included the thinning of trees by
poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
intensity was measured as the percentage of aboveground biomass (\%AGB)
lost between the first inventory in 1984 and five years after
disturbance (ref to be found) measured with the BIOMASS R package
\citep{Biomass2018}.

@@

The study site corresponds to a tropical rainforest with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH are mapped and measured annually since 1984. Trees are
first identified during inventories with a vernacular name assigned by
the field team, and afterward with a scientific name assigned by a
botanist during regular botanical campaigns. In 1984, specific
vernacular names are given to 62 commercial or common species whereas
more infrequent ones were identified under general identifiers only
distinguishing trees and palm trees. From 2003 botanical campaigns were
conducted every 5 to 6 years to identify all trees at the species level
but identification practices still varied among plots and campaigns.

@@

The functional approach used a dataset of 6 functional traits
representing leaf economics (leaves thickness, toughness, total
chlorophyll content and specific leaf area, the leaf area per unit dry
mass) and wood economics (wood specific gravity and bark thickness), and
life history traits (maximum specific height and seed mass). The trait
database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}
where trait values were assessed from a selection of individuals located
in nine permanent plots in French Guiana, including two in Paracou, and
comprised 294 botanical species pertaining to 157 botanical genera.
Missing trait values were filled using multivariate imputation by
chained equation (mice) from the mice R package \citep{Mice2011}.
Imputations were restricted within genus, or family when samples wre too
scarce, in order to account for the phylogenetic signal of the
functional traits. Whenever a species inventoried was not in the
dataset, it was attributed a set of traits values randomly sampled among
species of the same next higher taxonomic level (same genus or family).
As seed mass information corresponds to a classification into mass
classes, no data filling process was applied and analysis were
restricted to the 414 botanical species recorded.

All composition and diversity metrics corresponded to the average
obtained after 50 iterations of the taxonomic uncertainty propagation
framework and of the filling process of missing trait values.

@@

To counter taxonomic uncertainties due to the variability of botanical
identification protocols (see {[}\#protocols{]}), the taxonomic
composition and diversity analysis were conducted at the genus level,
\emph{i.e.} referring to the genus of observed or trialed botanical
names. Trajectories of communities taxonomic and functional variations
in composition after disturbance were followed in a two-dimensional
ordination space the 30 years monitored. Two NMDS were conducted to map
either taxonomic flora inventories or communities functional composition
based on the 7 leaf, stem and life history traits (without seed mass
classes). In both cases the NMDS were performed using occurrence-based
(Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures.
Trajectories along time in the plan were reported through the euclidean
distance of successive inventories to the reference inventories in 1989,
5 years after disturbance, when the uncertainty degree did not exceed
30\% of undetermined trees. The trajectories of the leaf and stem and
life traits were also visualized with the community weighted means
(CWM), representing the average trait value in a community weighted by
relative abundance of the species carrying each value
\citep{Diaz2007, Garnier2004}. To compensate the intrinsinc difference
among plots the trajectories corresponded to the differences along time
with the reference inventory in 1989. Species seed mass corresponded to
5 classes of increasing mass, seed mass trajectories were therefore
reported as the proportion of each class in the inventories.

>>>>

The experiment is a network of twelve 6.25ha plots that underwent a
gradient of three logging, thinning and fuelwood cutting treatments
(Table \ref{tab:Tab1}) according to a randomized plot design
with three replicate blocks of four plots. The disturbance corresponds
to averages of 10 trees removed per hectare with a diameter at 1.3 m
height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
(T3). Treatments T2 and T3 besides included the thinning of trees by
poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
intensity was measured as the percentage of aboveground biomass (\%AGB)
lost between the first inventory in 1984 and five years after
disturbance \citep{Piponiot2016} estimated with the BIOMASS R package
\citep{Biomass2018}.

@@

The study site corresponds to a tropical rainforest typical if the Guiana Shield with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH have been mapped and measured annually since 1984. Trees are
first identified during inventories with a vernacular name assigned by
the forest worker team, and afterward with a scientific name assigned by
botanists during regular botanical campaigns. In 1984, specific
vernacular names are given to 62 commercial or common species whereas
more infrequent ones were identified under general identifiers only
distinguishing trees and palm trees. From 2003, botanical campaigns have been
conducted every 5 to 6 years to identify all trees at the species level
but identification levels still varied among plots and campaigns.

@@

Six functional traits
representing leaf economics (leaves thickness, toughness, total
chlorophyll content and specific leaf area, the leaf area per unit dry
mass) and wood economics (wood specific gravity and bark thickness), and
life history traits (maximum specific height and seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were assessed from a selection of individuals located
in nine permanent plots in French Guiana, including two in Paracou, and
comprised 294 botanical species pertaining to 157 botanical genera.
Missing trait values were filled using multivariate imputation by
chained equation \citep{Mice2011}.
Imputations were restricted within genus, or family when samples were too
scarce, in order to account for the phylogenetic signal of the
functional traits. Whenever a species inventoried was not in the
dataset, it was attributed a set of traits values randomly sampled among
species of the same next higher taxonomic level (same genus or family).
As seed mass information corresponds to a classification into mass
classes, no data filling process was applied and analysis were
restricted to the 414 botanical species recorded.

All composition and diversity metrics were
obtained after 50 iterations of the taxonomic uncertainty propagation
framework and of the filling process of missing trait values.

@@

To counter taxonomic uncertainties due to the variability of botanical
identification levels (in space) and protocols (in time) (see {[}\#protocols{]}), the taxonomic
composition and diversity analysis were conducted at the genus level,
\emph{i.e.} referring to the genus of observed or trialed botanical
names. Trajectories of communities taxonomic and functional variations
in composition after disturbance were followed in a two-dimensional
ordination space the 30 years monitored. Two NMDS were conducted to map
either taxonomic flora inventories or communities functional composition
based on the 7 leaf, stem and life history traits (without seed mass
classes). In both cases the NMDS were performed using occurrence-based
(Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures.
Trajectories along time in the plan were reported through the euclidean
distance of successive inventories to the reference inventories in 1989,
5 years after disturbance, when the uncertainty degree did not exceed
30\% of undetermined trees. The trajectories of the leaf and stem and
life traits were also visualized with the community weighted means
(CWM), representing the average trait value in a community weighted by
relative abundance of the species carrying each value
\citep{Diaz2007, Garnier2004}. To compensate the intrinsinc difference
among plots the trajectories corresponded to the differences along time
with the reference inventory in 1989. Species seed mass corresponded to
5 classes of increasing mass, seed mass trajectories were therefore
reported as the proportion of each class in the inventories.
